Leadership across cultures and contexts: an exploratory analysis

### 2020-06-22 leadership_across_cultures_contexts.Rmd
Some edits to universal section in disucssion

We found strong evidence that several dimensions of leadership were universal across cultures and across contexts within cultures. Leaders are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures, and function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures (Figure \@ref(fig:varsupport)). Leaders benefited materially, reproductively, or socially in 50% or more of cultures (Figure \@ref(fig:varsupportcb)).

@@

Some common leadership dimensions were also relatively invariant across the four factors we examined: group context, subsistence strategy, continental region, and leader sex. These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous; and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal" we do not mean that all leaders have all these traits. Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group variation (i.e., the four factors). See Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport).

@@

The leadership features we derived by clustering leadership dimensions were represented across most cultures. Leaders performed three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions. Regarding the traits characteristic of leadership, leaders possessed some combination of individual capacities, many of which are necessarily prosocial, conforming to particular cultural norms, and receiving various benefits in exchange for their functions (Figure \@ref(fig:featuressupportplot)).

These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories that have drawn heavily on select ethnographic cases [@garfield_review_2019]. They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders. We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)). This feature corresponds, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2020].

The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for providing services, but that model does not emphasize the concrete material benefits seen here. The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, intelligence, conflict resolution, generosity, and status leveling mechanisms among largely egalitarian societies.

>>>>

We found strong evidence that several leadership dimensions were universal across cultures and across contexts within cultures. Leaders are seen as high status, knowledgeable or intelligent, and experienced or accomplished in about 80% or more of cultures, and function to resolve conflict, organize cooperation, and provide counsel or direction in over 70% of cultures (Figure \@ref(fig:varsupport)). Leaders benefited materially, reproductively, or socially in 50% or more of cultures (Figure \@ref(fig:varsupportcb)).

@@

Some common leadership dimensions were also relatively invariant across group context, subsistence strategy, continental region, and leader sex. These dimensions included qualities, such as being knowledgeable or intelligent, wealthy, strategically nepotistic, and generous; and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal" we do not mean that all leaders have all these traits. Instead, we mean that there is evidence for these traits across a solid majority of cultures that do not vary much by key measures of population or group variation (i.e., the four factors) (Table \@ref(tab:leaderuniversals) and Figure \@ref(fig:varsupport)).

@@

The leadership features we derived by clustering leadership dimensions were represented across most cultures. Leaders performed three broad functions: providing prosocial goods and services, strategizing, and organizing collective actions. Regarding characteristic traits, leaders possessed some combination of individual capacities, many of which are prosocial, conformed to particular cultural norms, and receivied various benefits (Figure \@ref(fig:featuressupportplot)).

These potentially universal dimensions of leadership correspond to several evolutionary theories that have drawn heavily on select ethnographic cases [@garfield_review_2019]. They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders. We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)). This feature corresponds, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2020].

The universal dimensions also partially correspond to the service-for-prestige model [@price_evolution_2014; @price_service-for-prestige_2015] in which leaders receive special status in exchange for providing services, but that model does not emphasize the concrete material benefits seen here. The absence of *High status* and *Aggressiveness* on this list of universals corresponds to the reverse dominance hierarchy model [@boehm_egalitarian_1993], which emphasizes knowledge, conflict resolution, generosity, and status leveling mechanisms among largely egalitarian societies.
